Cradles of the first stars: self-shielding, halo masses, and multiplicity

### 2019-11-25 paper.tex
Adding final revisions to paper and responses to referee.

Diffuse gas is optically thin to LW radiation, thus a background builds over time and can suppress \hh{} formation, delaying Pop III star formation. Furthermore, nearby sources of LW radiation can boost the intensity above the background value which facilitates further \hh{} dissociation in minihalos. This process can be counteracted with a sufficient amount of \hh{} already present within a halo, via \hh{} self-shielding. Halos with an \hh{} column density of N$_{\rm H2}$ $\geq$ 10$^{14}$ cm$^{-2}$ can suppress the photodissociation of \hh{} by LW photons \citep{Draine96}, giving the halo a chance to form Pop III stars. Other sources of the suppression of Pop III star formation are streaming baryonic velocities \citep{Tselia11, Greif11_Delay, Naoz12, OLeary12, Schauer19,Hirano17_Science}, arising from recombination, and dynamical heating, occurring during the virialization of halos \citep{Yoshida03, Fernandez14}. X-ray heating from high-mass X-ray binaries also has the potential to suppress Pop III star formation \citep{Jeon14}.

@@

For $\rm{H}_{0} = 70 \ \rm{km} \ \rm{s}^{-1} \ \rm{Mpc}^{-1}$, and assuming the Doppler broadening parameter, $b$, is equal to the circular velocity of the halo, Equation \ref{columnh2} and Equation \ref{shield} will give the shielding factor for a given core radius, $f_{c}$, redshift, and for a given \hh{} fraction halo profile. Figure \ref{fig:shield_mass} shows the shielding factor as a function of halo mass, for redshifts $z = 9$, $15$, and $25$ in red, blue and black, respectively. The solid lines indicate a core radius of $f_{c} = 10^{-2}$ and the dashed lines a core radius of $f_{c} = 10^{-4}$. To have an accurate depiction of the shielding factor for a given halo radius, the \hh{} fraction is given by the \hh{} fraction halo profiles at z = 24.1 from \citet{OShea08} (their Figure 13d) with $J_{21}$ = 0, which was extracted directly from their plot. These profiles were approximated with a piecewise power-law for simplicity \textbf{given by Equation \ref{h2fitting}}. 
\begin{equation} \label{h2fitting}
	log_{10}(f_{\rm H2}) = -0.482 \times log_{10}(R_{\rm vir}) - 4.714
\end{equation}
As the halo masses increase, the shielding factor increases, meaning there is less shielding occurring within that halo. This behavior occurs because at larger radii, there is less \hh{} available, and therefore, the effects of shielding are decreased. Across the top axis and plotted as a green line, the shielding factor as a function of \hh{} column density is shown for a circular velocity of 10 km/s. At each redshift, all halo masses, and for each core radius, we see the shielding factor is significantly smaller than one, implying that \hh{} shielding for these halos is almost at a maximum. The shielding factor will then significantly decrease the dissociation rate of \hh{}, allowing for the presence of \hh{} within a LW background. Smaller mass halos can then form Pop III stars since \hh{} will survive in the halo for a longer amount of time. We also see that for column densities at just under $10^{17} \rm cm^{-2}$, the shielding factor quickly drops to 0.1 for a halo with this particular circular velocity, showing that \hh{} shielding can affect the dissociation rate of \hh{} in relatively low halo masses.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{images/compare_JLW_mass.pdf}
    \caption{The halo mass versus the average LW intensity normalized by J21 versus for this work and a variety of other works. The black solid line is from M01 given our LW background, blue solid line from \citet{Trenti09_SFR} at $z=20$, the purple solid line from \citet{Visbal14} at $z=20$, blue open circle from \citet{Yoshida03}, green open squares from \citet{OShea08}, and red open triangles from \citet{Wise07_UVB}. J$_{LW}$ = 0 for the various sources is plotted here at J$_{\rm LW}$/J$_{21}$ = 10$^{-4}$. The solid bands of halo masses correspond to the characteristic masses of halos to contain a baryon fraction that is half of the cosmic mean for various redshifts and for the shown streaming velocities from \citet{Naoz13}. For each band, the maximum halo mass occurs at $z = 15$. For streaming velocities of $3.4 \sigma_{\rm vbc}$, $1.7 \sigma_{\rm vbc}$, and $1 \sigma_{\rm vbc}$, the minimum halo mass occurs at $z = 17, 22, \rm{and} \ 27$, respectively. \textbf{The minimum halo mass hosting cold dense gas for different streaming velocities in units of km/s from \citet{Schauer19} are shown as horizontal black lines with different linestyles.}}
    \label{fig:compare_JLW_mass}
\end{figure}

@@

We also do not include streaming velocities between baryons and dark matter within our simulation \citep{Tselia11, Greif11_Delay, Naoz12, OLeary12, Schauer19}. Streaming velocities can suppress star formation at very high ($z \ga 20$) redshifts only, because the streaming velocities decrease with time. Pop III star formation is delayed by an average of $\delta z = 4$ by increasing the halo mass needed to overcome the bulk velocity by about a factor of three \citep{Greif11_Delay}. The streaming velocities will generally result in lower gas densities within halos as well as an offset of the peak density from the center of the halo \citep{OLeary12}. \citet{Naoz13} estimated the minimum mass needed to retain the bulk of the baryons within the dark matter halo using a series of cosmological simulations, the results of which can be seen in Figure \ref{fig:compare_JLW_mass} (see shaded bands). These halos only reach up to about $2 - 3 \times 10^5 M_\odot$ when the streaming velocity is about $1\sigma_{\rm vbc}$, which is smaller than the star forming halos in our simulation. \citet{Naoz13} mention that at the largest streaming velocity, their simulations do not match expected values due to a small sample size of high mass halos. \textbf{A more recent study by \citet{Schauer19}, who used cosmological simulations to study cold gas content within halos for varying streaming velocities, showed that high streaming velocities restrict first star formation to atomic cooling halos. Their minimum halo masses which host cold dense gas for various streaming velocities are shown as black horizontal lines with different linestyles in Figure \ref{fig:compare_JLW_mass}.} Our results indicate that halos collapse at larger masses that are not as affected by typical streaming velocities ($<1\sigma_{\rm vbc}$).  In the evolution of such a halo, it would have started to accumulate gas above these critical masses while also being irradiated by a LW background.  Afterwards, the collapse will be controlled by the processes discussed in Section \ref{sec:variations}, and thus streaming velocities would likely not affect our results.

>>>>

Diffuse gas is optically thin to LW radiation, thus a background builds over time and can suppress \hh{} formation, delaying Pop III star formation. Furthermore, nearby sources of LW radiation can boost the intensity above the background value which facilitates further \hh{} dissociation in minihalos. This process can be counteracted with a sufficient amount of \hh{} already present within a halo, via \hh{} self-shielding. Halos with an \hh{} column density of N$_{\rm H2}$ $\geq$ 10$^{14}$ cm$^{-2}$ can suppress the photodissociation of \hh{} by LW photons \citep{Draine96}, giving the halo a chance to form Pop III stars. Other sources of the suppression of Pop III star formation are streaming baryonic velocities \citep{Tselia11, Greif11_Delay, Naoz12, OLeary12, Schauer19, Hirano17_Science}, arising from recombination, and dynamical heating, occurring during the virialization of halos \citep{Yoshida03, Fernandez14}. X-ray heating from high-mass X-ray binaries also has the potential to suppress Pop III star formation \citep{Jeon14}.

@@

For $\rm{H}_{0} = 70 \ \rm{km} \ \rm{s}^{-1} \ \rm{Mpc}^{-1}$, and assuming the Doppler broadening parameter, $b$, is equal to the circular velocity of the halo, Equation \ref{columnh2} and Equation \ref{shield} will give the shielding factor for a given core radius, $f_{c}$, redshift, and for a given \hh{} fraction halo profile. Figure \ref{fig:shield_mass} shows the shielding factor as a function of halo mass, for redshifts $z = 9$, $15$, and $25$ in red, blue and black, respectively. The solid lines indicate a core radius of $f_{c} = 10^{-2}$ and the dashed lines a core radius of $f_{c} = 10^{-4}$. To have an accurate depiction of the shielding factor for a given halo radius, the \hh{} fraction is given by the \hh{} fraction halo profiles at z = 24.1 from \citet{OShea08} (their Figure 13d) with $J_{21}$ = 0, which was extracted directly from their plot. These profiles were approximated with a piecewise power-law for simplicity \textbf{given by $\log_{10}(f_{\rm H2}) = -0.482 \times \log_{10}(R_{\rm vir}) - 4.714$}. 
As the halo masses increase, the shielding factor increases, meaning there is less shielding occurring within that halo. This behavior occurs because at larger radii, there is less \hh{} available, and therefore, the effects of shielding are decreased. Across the top axis and plotted as a green line, the shielding factor as a function of \hh{} column density is shown for a circular velocity of 10 km/s. At each redshift, all halo masses, and for each core radius, we see the shielding factor is significantly smaller than one, implying that \hh{} shielding for these halos is almost at a maximum. The shielding factor will then significantly decrease the dissociation rate of \hh{}, allowing for the presence of \hh{} within a LW background. Smaller mass halos can then form Pop III stars since \hh{} will survive in the halo for a longer amount of time. We also see that for column densities at just under $10^{17} \rm cm^{-2}$, the shielding factor quickly drops to 0.1 for a halo with this particular circular velocity, showing that \hh{} shielding can affect the dissociation rate of \hh{} in relatively low halo masses.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{images/compare_JLW_mass.pdf}
    \caption{The halo mass versus the average LW intensity normalized by J21 versus for this work and a variety of other works. The black solid line is from M01 given our LW background, blue solid line from \citet{Trenti09_SFR} at $z=20$, the purple solid line from \citet{Visbal14} at $z=20$, blue open circle from \citet{Yoshida03}, green open squares from \citet{OShea08}, and red open triangles from \citet{Wise07_UVB}. J$_{LW}$ = 0 for the various sources is plotted here at J$_{\rm LW}$/J$_{21}$ = 10$^{-4}$. \textbf{The left panel represents halo masses corresponding to different streaming velocities from \citet{Naoz13} and \citet{Schauer19}.}The solid bands of halo masses correspond to the characteristic masses of halos to contain a baryon fraction that is half of the cosmic mean for various redshifts and for the shown streaming velocities from \citet{Naoz13}. For each band, the maximum halo mass occurs at $z = 15$. For streaming velocities of $3.4 \sigma_{\rm vbc}$, $1.7 \sigma_{\rm vbc}$, and $1 \sigma_{\rm vbc}$, the minimum halo mass occurs at $z = 17, 22, \rm{and} \ 27$, respectively. \textbf{The minimum halo mass hosting cold dense gas for different streaming velocities from \citet{Schauer19} are shown as horizontal black lines.}}
    \label{fig:compare_JLW_mass}
\end{figure}

@@

We also do not include streaming velocities between baryons and dark matter within our simulation \citep{Tselia11, Greif11_Delay, Naoz12, OLeary12, Schauer19, Hirano17_Science}. Streaming velocities can suppress star formation at very high ($z \ga 20$) redshifts only, because the streaming velocities decrease with time. Pop III star formation is delayed by an average of $\delta z = 4$ by increasing the halo mass needed to overcome the bulk velocity by about a factor of three \citep{Greif11_Delay}. The streaming velocities will generally result in lower gas densities within halos as well as an offset of the peak density from the center of the halo \citep{OLeary12}. \citet{Naoz13} estimated the minimum mass needed to retain the bulk of the baryons within the dark matter halo using a series of cosmological simulations, the results of which can be seen in Figure \ref{fig:compare_JLW_mass} (see shaded bands). These halos only reach up to about $2 - 3 \times 10^5 M_\odot$ when the streaming velocity is about $1\sigma_{\rm vbc}$, which is smaller than the star forming halos in our simulation. \citet{Naoz13} mention that at the largest streaming velocity, their simulations do not match expected values due to a small sample size of high mass halos.

\textbf{A more recent study by \citet{Schauer19}, who used cosmological simulations to study cold gas content within halos for varying streaming velocities, showed that high streaming velocities restrict first star formation to atomic cooling halos. Their minimum halo masses which host cold dense gas for various streaming velocities are shown as black horizontal lines with different linestyles in Figure \ref{fig:compare_JLW_mass}. \citet{Hirano17_Science} investigated the effect that streaming velocities have on black hole formation in the early universe and find that in regions with high streaming velocities and for z $> 30$, first generation star formation can be suppressed in minihalos up to the atomic cooling limit, and may produce massive black hole seeds.}

Our results indicate that halos collapse at larger masses that are not as affected by typical streaming velocities ($<1\sigma_{\rm vbc}$). \textbf{We also find that Pop III stars begin to form at much later times than the halos found in \citet{Hirano17_Science}.} In the evolution of such a halo, it would have started to accumulate gas above these critical masses while also being irradiated by a LW background.  Afterwards, the collapse will be controlled by the processes discussed in Section \ref{sec:variations}, and thus streaming velocities would likely not affect our results.

### 2019-11-25 referee_2.tex
Adding final revisions to paper and responses to referee.

\begin{referee}
    A -- In this paper, some very long paragraphs can be found, especially in Sections 3 and 4. Please break these long paragraphs down in smaller ones to increase the read flow. 
\end{referee}
    These paragraphs have been broken down.

@@

    B -- The convergence study is a great addition to the paper. However, the lowest limit explored here (0.625pc) is a factor of roughly 100 larger than the scale on which fragmentation can be seen (compare eg the 1000AU regions, in which several Pop III stars form in work by Stacy+16). I recommend to explicitly state down to which resolution the convergence was tested in Section 2.2.1 and adding these limitations to the "Caveats" section. Instead of single Pop III stars, this paper is a study of the multiplicity of Pop III star formation regions. Please make this clear in the paper. 
\end{referee}
    The lowest limit has been added to section 2.2.1.

@@

    C -- - Is the SFRD really in physical (and not comoving) units (Figure 4 / Section 3)? Please provide values for the comparison papers (Xu+13, Magg+16) and include a comparison to Jaacks+19. 
\end{referee}

@@

\begin{referee}
    D -- - The initialization at z=130 should be mentioned as a caveat. 
\end{referee}
	We have added the following text to Section 2.1 after stating the initialization redshift, ``We note that this initialization redshift is too low, even with 2nd order Lagrangian perturbation theory, for a small cosmological volume and results in delayed structure formation at very high redshift $z > 20$.''

\begin{referee}
    E -- - The analytic calculation of the self-shielding of the halo now takes into account a realistic halo profile. However, there are three issues: 
    E.1 -- -- In the first paragraph of Section 2.4, it still reads "with a constant H2 fraction". Please correct this. 
\end{referee}
    This has been corrected.

\begin{referee}
    E.2 -- -- For completeness, please provide the readers with the H2 radial fit to the O'Shea \& Normal 07 paper that was performed. 
\end{referee}
    The citation previously written was referencing their first paper in their series of Pop III star formation. It should have referenced their second paper in the series. This has been corrected.

@@

\begin{referee}
    E.3 -- -- Compared to the previous version with a constant profile, the shielding factor increases for higher halo masses (in the previous version of the paper, there was a slight decrease): leading to less shielding. I do not understand where this is coming from, a more massive halo has a larger overall gas mass, and therefore more H2 and more shielding. Maybe it has to do with the scaling of the H2 fraction?  It is important to know how the authors scale the H2 fraction as a function of radius for different halo masses (see point above). If this is not the reason, please explain what else is causing more massive haloes to be worse at self-shielding.  
\end{referee}
    The shielding factor increases for more massive halos because these halos have larger virial radii, at which point there is significantly less \hh{} compared to smaller radii. So halos with smaller masses, and thus smaller virial radii, have more \hh{} built up at the edges compared to their larger counterparts, resulting in a more effective shield at the virial radius. We have clarified this point in the text.

\begin{referee}
    F -- - Thank you for adding the global LWBG as a dashed line in Figure 3. Please also mention at some point in the text or add a legend that this is again given in units of J21. 
\end{referee}
    The solid red line in the legend refers to both the solid and the dashed red lines. This detail has been added to the caption and to the text.

\begin{referee}
    G -- - I see that retrospectively calculating the time-averaged LWBG is not possible without repeating the study. However, please mention the possibility for a correlation between the time-integrated LWBG and the host halo mass in Section 3.2 and summary and be more explicit that this is the LWBG at star formation (eg in the introduction and in Section 3.2 and in the second bullet point of the Conclusions).  
\end{referee}
    The possible correlation between the LW background and the host halo mass has been added to the requested sections.

@@

\begin{referee}
    H -- - The authors put their work in comparison to Schauer+19, ApJL 877, 5 who match the star formation efficiency in minihaloes to the observed 21cm signal. They should rather compare to Schauer+19 MNRAS, 484, 3510 who performed cosmological hydrodynamical simulations for the minimum and average halo mass in different streaming velocity regions. Please include their limits in your Figure 10, as done with the Naoz+13 results, and a short comparison to Hirano+18 (their Science paper). Please include these references also in the introduction. Also the sentence "Often in the literature, streaming velocities are only studied for a few haloes" seems outdated.
\end{referee}
	The limits from Schauer+19 have been added to Figure 10, and the previous description regarding the 21cm signal has been removed. 
	--conclusions from this?
	--weird to me how the units of the streaming velocities are different. Change this?
	--unbolded addition to citation in caveat section

	--need to add the Hirano+18 science paper to caveats section

	Schauer+19 and Hirano+18 have been added to the introduction (unbolded -- fix this?).

>>>>

\begin{referee}
    In this paper, some very long paragraphs can be found, especially in Sections 3 and 4. Please break these long paragraphs down in smaller ones to increase the read flow. 
\end{referee}
    These paragraphs have been broken down.

@@

    The convergence study is a great addition to the paper. However, the lowest limit explored here (0.625pc) is a factor of roughly 100 larger than the scale on which fragmentation can be seen (compare eg the 1000AU regions, in which several Pop III stars form in work by Stacy+16). I recommend to explicitly state down to which resolution the convergence was tested in Section 2.2.1 and adding these limitations to the "Caveats" section. Instead of single Pop III stars, this paper is a study of the multiplicity of Pop III star formation regions. Please make this clear in the paper. 
\end{referee}
    The lowest limit has been added to section 2.2.1.

@@

    - Is the SFRD really in physical (and not comoving) units (Figure 4 / Section 3)? Please provide values for the comparison papers (Xu+13, Magg+16) and include a comparison to Jaacks+19. 
\end{referee}

@@

\begin{referee}
    - The initialization at z=130 should be mentioned as a caveat. 
\end{referee}
	We have added the following text to Section 2.1 after stating the initialization redshift, ``We note that this initialization redshift is too low, even with 2nd order Lagrangian perturbation theory, for a small cosmological volume and results in delayed structure formation at very high redshift $z > 20$.''

\begin{referee}
    - The analytic calculation of the self-shielding of the halo now takes into account a realistic halo profile. However, there are three issues: 
    -- In the first paragraph of Section 2.4, it still reads "with a constant H2 fraction". Please correct this. 
\end{referee}
    This has been corrected.

\begin{referee}
    -- For completeness, please provide the readers with the H2 radial fit to the O'Shea \& Normal 07 paper that was performed. 
\end{referee}
    The citation previously written was referencing their first paper in their series of Pop III star formation. It should have referenced their second paper in the series. This has been corrected.

@@

\begin{referee}
    -- Compared to the previous version with a constant profile, the shielding factor increases for higher halo masses (in the previous version of the paper, there was a slight decrease): leading to less shielding. I do not understand where this is coming from, a more massive halo has a larger overall gas mass, and therefore more H2 and more shielding. Maybe it has to do with the scaling of the H2 fraction?  It is important to know how the authors scale the H2 fraction as a function of radius for different halo masses (see point above). If this is not the reason, please explain what else is causing more massive haloes to be worse at self-shielding.  
\end{referee}
    The shielding factor increases for more massive halos because these halos have larger virial radii, at which point there is significantly less \hh{} compared to smaller radii. So halos with smaller masses, and thus smaller virial radii, have more \hh{} built up at the edges compared to their larger counterparts, resulting in a more effective shield at the virial radius. We have clarified this point in the text.

\begin{referee}
    - Thank you for adding the global LWBG as a dashed line in Figure 3. Please also mention at some point in the text or add a legend that this is again given in units of J21. 
\end{referee}
    The solid red line in the legend refers to both the solid and the dashed red lines. This detail has been added to the caption and to the text.

\begin{referee}
    - I see that retrospectively calculating the time-averaged LWBG is not possible without repeating the study. However, please mention the possibility for a correlation between the time-integrated LWBG and the host halo mass in Section 3.2 and summary and be more explicit that this is the LWBG at star formation (eg in the introduction and in Section 3.2 and in the second bullet point of the Conclusions).  
\end{referee}
    The possible correlation between the LW background and the host halo mass has been added to the requested sections.

@@

\begin{referee}
    - The authors put their work in comparison to Schauer+19, ApJL 877, 5 who match the star formation efficiency in minihaloes to the observed 21cm signal. They should rather compare to Schauer+19 MNRAS, 484, 3510 who performed cosmological hydrodynamical simulations for the minimum and average halo mass in different streaming velocity regions. Please include their limits in your Figure 10, as done with the Naoz+13 results, and a short comparison to Hirano+18 (their Science paper). Please include these references also in the introduction. Also the sentence "Often in the literature, streaming velocities are only studied for a few haloes" seems outdated.
\end{referee}
    We have added a panel to the left of Figure 10 to de-clutter the main figure and to separate the halo mass ranges and minimums related to streaming velocitys.

    The limits from Schauer+19 have been added to the left panel of Figure 10, and the previous description regarding the 21cm signal has been removed. Their citation has been added to the list of citations investigating streaming velocities.

    We have added a short compariosn to Hirano+18 and have also added their citation to the list of citations mentioned previously.

    These citations have also been added to the introduction.
